---
title: "ncaahoopR EDA"
output:
  pdf_document: default
  html_notebook: default
---

```{r}
# install.packages("devtools")
devtools::install_github("lbenz730/ncaahoopR", force = TRUE)
```

```{r}
library(tidyverse)
library(backports)
library(ncaahoopR)
```

```{r}
data(ids)
view(ids)
```

```{r}
ark_pbp <- get_pbp("Arkansas", season = "2022-23")
```

```{r}
ark_roster <- get_roster("Arkansas", season = "2022-23")
ark_schedule <- get_schedule("Arkansas", season = "2022-23")
```

```{r}
wp_chart(game_id = 401484594, home_col = "black", away_col = "red")
```

```{r}
gg_wp_chart(game_id = 401484594, home_col = "black", away_col = "red", show_labels = T)
```

```{r}
game_flow(game_id = 401484594, home_col = "black", away_col = "red")
```

```{r}
average_win_prob(game_id = 401484594, include_spread = T)
```
ESPN game_id for which to compute time-based average win probability (from perspective of home team).


```{r}
average_score_diff(game_id = 401484594)
```

ESPN game_id for which to compute time-based average score differential (from perspective of home team).


```{r}
# switched to the LSU game because the data for the South Carolina game was not there yet
get_shot_locs(game_id = 401484577)
```
Returns data frame with shot location data when available.


```{r}
#' Plot all shots from one game
#'
#' Plots game shot locations from one game.
#'
#' @param game_id ESPN game_id for which to render shot chart
#' @param heatmap Logical, Use a density-type heatmap, Default = False
#' @export
#'
#'
game_shot_chart <- function(game_id, heatmap = F){
  if(any(is.na(game_id))) {
    error("game_id missing with no default")
  }
  if(length(game_id) > 1) {
    error("game_shot_chart only takes in a single game_id")
  }
  
  shot_loc_df <- get_shot_locs(game_id)
  if(!is.null(shot_loc_df)) {
    teams <- sort(unique(shot_loc_df$team_name))
    game_title <- paste0(teams[1]," vs. ", teams[2])
    color <- c(shot_loc_df$color[shot_loc_df$team_name == teams[1]][1],
               shot_loc_df$color[shot_loc_df$team_name == teams[2]][1])
    
    date <- format(as.Date(shot_loc_df$date[1]), "%B %d, %Y")
    
    if(heatmap){
      full_title <- paste0(game_title," \n ",date)
      title <- cowplot::ggdraw() +
        cowplot::draw_label(full_title,
                            fontface = 'bold')
      p <- team_shot_chart(game_id, teams[1], heatmap = T) +
        theme(legend.position = "bottom")
      p2 <- team_shot_chart(game_id, teams[2], heatmap=T) +
        labs(caption="") +
        theme(legend.position = "bottom")
      p3 <- cowplot::plot_grid(p,p2)
      p1 <- cowplot::plot_grid(title, p3, ncol = 1, rel_heights = c(0.1, 1))
      return(p1)
    }
    
    p1 <- suppressMessages(ggplot2::ggplot() +
                             ggplot2::geom_point(data = shot_loc_df,
                                                 aes(
                                                   x = x,
                                                   y = y,
                                                   shape = team_name,
                                                   color = outcome),
                                                 size = 3) +
                             ggplot2::geom_polygon(data = court, aes(x = x, y = y, group = group), col = "gray") +
                             ggplot2::geom_point(alpha = 0.2, size = 1.5) +
                             ggplot2::scale_color_manual(values = color) +
                             ggplot2::xlab("") +
                             ggplot2::ylab("")  +
                             ggplot2::coord_flip() +
                             ggplot2::scale_x_reverse() +
                             ggplot2::theme_void() +
                             ggplot2::theme(
                               axis.text.x = element_blank(),
                               axis.text.y = element_blank(),
                               axis.ticks.x = element_blank(),
                               axis.ticks.y = element_blank(),
                               axis.title = element_blank(),
                               legend.position = "bottom",
                               legend.direction = 'vertical',
                               plot.title = element_text(size = 16, hjust = 0.5),
                               plot.subtitle = element_text(size = 12, hjust = 0.5),
                               plot.caption = element_text(size = 8, hjust = 0),
                               plot.background = element_rect(fill = 'cornsilk')) +
                             ggplot2::labs(
                               title = game_title,
                               subtitle = date,
                               color = 'Made',
                               shape = 'Team',
                               caption = "Jacob Haarala (@jhaarala23) Data Accessed via ncaahoopR"))
    return(p1)
  }
}
```



```{r}
#' Plot all shots for a team
#'
#' Plots team shot locations from one game or multiple games.
#'
#' @param game_ids Vector of ESPN game_ids
#' @param team Team Name
#' @param heatmap Use a density-type heatmap (Default = FALSE)
#' @export
team_shot_chart <- function(game_ids, team, heatmap = F) {
  if(any(is.na(game_ids))) {
    error("game_ids missing with no default")
  }
  if(any(is.na(team))) {
    error("team missing with no default")
  }
  df <- get_shot_locs(game_ids)
  
  if(!is.null(df)) {
    side_one <- court %>% filter(side == 1)
    team_shots <- df %>% filter(team_name %in% c(team, dict$ESPN_PBP[dict$ESPN == team]))
    
    ### flip shots if they are on the wrong side
    team_shots[team_shots$y > 47, "x"] <- 50 - team_shots[team_shots$y > 47, "x"]
    team_shots[team_shots$y > 47, "y"] <- 94 - team_shots[team_shots$y > 47, "y"]
    
    ### only pick one color
    color <- as.character(unique(team_shots$color))[1]
    
    if(heatmap){
      p1 <-
        ggplot2::ggplot() +
        ggplot2::stat_density_2d(data = team_shots,
                                 aes(x = x, y = y, fill = stat(density / max(density))),
                                 geom = "raster", contour = FALSE, interpolate = TRUE, n = 200) +
        ggplot2::geom_polygon(data = side_one, aes(x = x, y = y, group = group), col = "gray") +
        ggplot2::geom_point(alpha = 0.2, size = 1.5) +
        ggplot2::coord_equal() +
        ggplot2::xlab("") +
        ggplot2::ylab("")  +
        ggplot2::scale_fill_viridis_c("Shot Frequency",
                                      limits = c(0, 1),
                                      breaks = c(0, 1),
                                      labels = c("Lower", "Higher"),
                                      option = "plasma") +
        ggplot2::theme_void() +
        ggplot2::theme(
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.x = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title = element_blank(),
          plot.title = element_text(size = 16, hjust = 0.5),
          plot.subtitle = element_text(size = 12, hjust = 0.5),
          plot.caption = element_text(size = 8, hjust = 0)) +
        ggplot2::labs(
          title = paste0(team," Shots"),
          shape = "Shot Outcome",
          caption = "Meyappan Subbaiah (@msubbaiah1) Data Accessed via ncaahoopR")
      return(p1)
    }
    
    p1 <-
      suppressMessages(ggplot2::ggplot() +
                         ggplot2::geom_point(data = team_shots,
                                             aes(
                                               x = x,
                                               y = y,
                                               shape = team_name),
                                             color = outcome,
                                             size = 3) +
                         ggplot2::geom_polygon(data = side_one, aes(x = x, y = y, group = group), col = "gray") +
                         ggplot2::geom_point(alpha = 0.2, size = 1.5) +
                         ggplot2::scale_color_manual(values = color) +
                         ggplot2:: xlab("") +
                         ggplot2::ylab("")  +
                         ggplot2::theme_void() +
                         ggplot2::theme(
                           axis.text.x = element_blank(),
                           axis.text.y = element_blank(),
                           axis.ticks.x = element_blank(),
                           axis.ticks.y = element_blank(),
                           axis.title = element_blank(),
                           plot.title = element_text(size = 16, hjust = 0.5),
                           plot.subtitle = element_text(size = 12, hjust = 0.5),
                           plot.caption = element_text(size = 8, hjust = 0),
                           plot.background = element_rect(fill = 'cornsilk')) +
                         ggplot2::labs(
                           title = paste(team, "Shots"),
                           shape = "Shot Outcome",
                           caption = "Jacob Haarala (@jhaarala23) Data Accessed via ncaahoopR"))
    return(p1)
  }
}
```



```{r}
game_shot_chart(game_id = 401484577, heatmap = F)
```




```{r}
team_shot_chart(game_id = 401484577, team = "Arkansas", heatmap = T)
```

If you want to do this over a span of games you can put multiple game ids in a vector for the 'game_id' parameter.


